---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading libraries
```{r loading_libraries}
library(plyr)
library(dplyr)
library(ggplot2)
```

```{r}
knitr::opts_chunk$set(fig.path = 'figure/')
```

## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv")
str(activity)
```

## What is mean total number of steps taken per day?
### Histogram of the total number of steps taken each day
```{r histogram}
activity_by_date <- group_by(activity, date)
activity_sum_by_date <- summarise(activity_by_date, daily_steps = sum(steps, na.rm = TRUE))
hist(activity_sum_by_date$daily_steps, breaks = 10, xlab = "Daily Steps", main = "Histogram of Daily Steps")
```

### Mean and median total number of steps taken per day
```{r}
mean_steps <- mean(activity_sum_by_date$daily_steps)
median_steps <- median(activity_sum_by_date$daily_steps)
```
The mean steps taken per day are `r mean_steps`.

The median steps taken per day are `r median_steps`.

## What is the average daily activity pattern?

### Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r daily_activity_pattern}
activity_by_interval <- group_by(activity, interval)
activity_mean_by_interval <- summarise(activity_by_interval, mean_steps = mean(steps, na.rm = TRUE))
plot(activity_mean_by_interval$interval, activity_mean_by_interval$mean_steps, type = "l", xlab = "Interval", ylab = "Mean Steps", main = "Average steps during a day")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_steps_interval <- filter(activity_mean_by_interval, mean_steps == max(mean_steps))
max_interval <- max_steps_interval$interval[1]
max_steps <- max_steps_interval$mean_steps[1]
```
The 5-minute interval `r max_interval` has the maximum average steps of `r max_steps` steps.

## Imputing missing values
### Rows with NAs
```{r}
activityWithNa <- filter(activity, is.na(steps) | is.na(date) | is.na(interval))
numberOfRowsWithNa <- dim(activityWithNa)[1]
```
There are `r numberOfRowsWithNa` rows with NA values.

### Filling missing values
Let's assume that we are very strict, so we set all NA values to 0.
```{r}
activity_imputed <- mutate(activity, steps_imputed = steps %>% mapvalues(NA, 0))
head(activity_imputed)
```

### Histogram with imputed values
```{r histogram_imputed}
activity_imputed_by_date <- group_by(activity_imputed, date)
activity_imputed_sum_by_date <- summarise(activity_imputed_by_date, daily_steps = sum(steps_imputed))
hist(activity_imputed_sum_by_date$daily_steps, breaks = 10, xlab = "Daily Steps Imputed", main = "Histogram of Daily Steps Imputed")
```

### Mean and median total number of steps taken per day
```{r}
mean_steps_imputed <- mean(activity_imputed_sum_by_date$daily_steps)
median_steps_imputed <- median(activity_imputed_sum_by_date$daily_steps)
```
The mean steps taken per day are `r mean_steps_imputed` (without imputing values it was `r mean_steps`).

The median steps taken per day are `r median_steps_imputed` (without imputing values it was `r median_steps`).

## Are there differences in activity patterns between weekdays and weekends?
Add a real date column and calculate if the date is a weekend day or a weekday day.
```{r}
activity_imputed_date <- mutate(activity_imputed, date_Date = as.Date(date), weekday = weekdays(date_Date, abbreviate = TRUE), weekend = factor(1 * (weekday %in% c("Sa", "So")), labels = c("weekday", "weekend")))
head(activity_imputed_date)
```

Now look at the difference between weekdays and weekend.
```{r weekend_pattern}
activity_imputed_by_interval <- group_by(activity_imputed_date, interval, weekend)
activity_imputed_mean_by_interval <- summarise(activity_imputed_by_interval, mean_steps = mean(steps, na.rm = TRUE))
g <- ggplot(activity_imputed_mean_by_interval, aes(interval, mean_steps))
g + geom_line() + facet_grid(weekend ~ .)
```
